---
title: "orthopedic_surgery"
author: "LEE CHUNGGAK"
date: '2021 3 25 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(tidyverse)
library(readxl)
library(car)

setwd("C:/Users/Administrator/work/1_통계자문/orthopedic_surgery")

v <- c("post_er_at_side", "post_er_90", "post_ir_90", "post_ir_back",
       "post_er_strength", "post_ir_strength")

d <- read_excel("OS 이용걸교수님 논문통계데이터.xlsx",
                col_types = "text",
                skip = 1) %>% 
    rename(post_er_at_side = 27, post_er_90 = 28, post_ir_90 = 29,
           post_ir_back = 31, post_er_strength = 32, post_ir_strength = 33,
           sbs = 40) %>% 
    select(sbs, all_of(v)) %>% 
    mutate(sbs = as.factor(as.integer(sbs)),
           post_er_at_side = as.integer(post_er_at_side),
           post_er_90 = as.integer(post_er_90),
           post_ir_90 = as.integer(post_ir_90),
           post_ir_back = as.integer(post_ir_back),
           post_er_strength = as.double(post_er_strength),
           post_ir_strength = as.double(post_ir_strength)) %>% 
    drop_na(sbs)
```

## Post ER at Side
```{r include = FALSE}
i <- "post_er_at_side"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```

## Post ER 90'
```{r include = FALSE}
i <- "post_er_90"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```

## Post IR 90'
```{r include = FALSE}
i <- "post_ir_90"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```

## Post IR Back'
```{r include = FALSE}
i <- "post_ir_back"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```

## Post ER Strength
```{r include = FALSE}
i <- "post_er_strength"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```

## Post IR Strength
```{r include = FALSE}
i <- "post_ir_strength"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```