---
title: "orthopedic_surgery"
author: "LEE CHUNGGAK"
date: '2021 4 6 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(tidyverse)
library(readxl)
library(car)

setwd("C:/Users/Administrator/work/1_임상의학연구소/1_통계/orthopedic_surgery")

v <- c("toilet", "post_vas_rest", "post_vas_rom", "pain", "func",
       "active_ff", "strength_er", "satisfaction", "total")

d <- read_excel("data/21.03.29 수정 SBS 논문 data.xlsx",
                col_types = "text",
                skip = 1) %>% 
    rename(sbs = 40, toilet = 42, post_vas_rest = 43, post_vas_rom = 44,
           func = 46,
           active_ff = 47, strength_er = 48, total = 50) %>% 
    select(sbs, all_of(v)) %>% 
    mutate_all(as.integer) %>% 
    mutate(sbs = as.character(sbs),
           toilet = as.character(toilet),
           sum_pain_func = pain + func) %>% 
    drop_na(sbs)
```

## Toilet

Cross table

```{r echo = FALSE}
t <- table(d$sbs, d$toilet)
print(t)
print(prop.table(t))
```

Chi-square test

```{r echo = FALSE}
print(chisq.test(t))
```



## Post VAS Rest

```{r include = FALSE}
i <- "post_vas_rest"
fml <- as.formula(paste(i, "~ sbs"))
```

Jitter plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_jitter()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```



## Post VAS Rom

```{r include = FALSE}
i <- "post_vas_rom"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```



## Sum of Pain and Function

```{r include = FALSE}
i <- "sum_pain_func"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```



## Pain

```{r include = FALSE}
i <- "pain"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```



## Function

```{r include = FALSE}
i <- "func"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```



## Satisfaction

```{r include = FALSE}
i <- "satisfaction"
fml <- as.formula(paste(i, "~ sbs"))
```

Box plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(x = "sbs", y = i)) +
        geom_boxplot()
```

Quantile-quantile plot

```{r echo = FALSE, warning = FALSE}
ggplot(data = d, aes_string(sample = i)) +
        geom_qq() +
        geom_qq_line() + 
        annotate(geom = "text", x = -Inf, y = Inf, hjust = 0, vjust = 1,
                 label = paste0("Shapiro test p-value: ", sprintf("%0.4f", shapiro.test(pull(d, i))[["p.value"]])))
```

Levene test

```{r echo = FALSE}
leveneTest(fml, data = d)
```

Bartlett test

```{r echo = FALSE}
bartlett.test(fml, data = d)
```

AOV

```{r echo = FALSE}
summary(aov(fml, data = d))
```